---
title: 'Coursera Data Science Project: Statistical Inference'
author: "Rob Rolleston"
date: "August 23, 2015"
output: pdf_document
---

###  Introduction
This is the project for the statistical inference class. In it, I will use simulation to explore inference and do some simple inferential data analysis. The project consists of two parts:

1. A simulation exercise.
2. Basic inferential data analysis.

## Simulation Excercise
tbd

### Simulations
```{r}
nosim <- 1000
n <- 40
lambda <- 0.2
populationMean <- 1./lambda
populationSD <- 1./lambda
sampleData <- matrix(rexp(nosim * n, rate=lambda), nosim)
```

### Sample Mean versus Theoretical Mean
```{r}
sampleMeans <- apply(sampleData, 1, mean)
sampleMean <- mean(sampleMeans)
sampleSD <- sd(sampleMeans)

inferredSD <- populationSD/sqrt(n)
```
```{r, echo=FALSE}
hist(sampleMeans)
```

### Sample Variance versus Theoretical Variance
tbd

### Distribution
tbd
```{r}
x <- seq(0,50, length=100)
hx <- dexp(x, rate=lambda)
```
```{r, echo=FALSE}
plot(x, hx, type="l")
```

***

##  Basic Inferential Data Analysis
tbd

### Load and explore data
tbd
```{r}
library(datasets)
toothGrowth <- ToothGrowth
```

### Data Summary
tbd

### Compare tooth growth by supp and dose
tbd

### Conclusions
tbd



