---
title: 'Coursera Data Science Project: Statistical Inference (Part 2)'
author: "Rob Rolleston"
date: "August 23, 2015"
output: pdf_document
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=3, fig.path='figure/',
                      echo=TRUE, warning=FALSE, message=FALSE)
```
```{r echo=FALSE}
library(ggplot2, quietly = TRUE, verbose=FALSE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, verbose=FALSE, warn.conflicts = FALSE)
library(tidyr, quietly = TRUE, verbose=FALSE, warn.conflicts = FALSE)
```
###  Introduction
This is the project for the statistical inference class. In it, I will use simulation to explore inference and do some simple inferential data analysis. The project consists of two parts:

1. A simulation exercise.
2. Basic inferential data analysis (this report)


##  Basic Inferential Data Analysis
tbd

### Load and explore data
tbd
```{r}
library(datasets)
data("ToothGrowth")
```

### Data Summary
tbd

### Compare tooth growth by supp and dose
tbd

### Conclusions
tbd



