---
title: 'Coursera Data Science Project: Statistical Inference (Part 1)'
author: "Rob Rolleston"
date: "August 23, 2015"
output: pdf_document
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=3, fig.path='figure/',
                      echo=TRUE, warning=FALSE, message=FALSE)
```
```{r echo=FALSE}
library(ggplot2, quietly = TRUE,verbose=FALSE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, verbose=FALSE, warn.conflicts = FALSE)
library(tidyr, quietly = TRUE, verbose=FALSE, warn.conflicts = FALSE)
```
###  Introduction
This is the project for the statistical inference class. In it, I will use simulation to explore inference and do some simple inferential data analysis. The project consists of two parts:

1. A simulation exercise. (this report)
2. Basic inferential data analysis.

## Simulation Excercise
tbd

### Simulations
```{r}
nosim <- 1000
n <- 40
lambda <- 0.2
populationMean <- 1./lambda
populationSD <- 1./lambda
sampleData <- matrix(rexp(nosim * n, rate=lambda), nosim)
hist(sampleData)

sampleData_tbl <- gather(tbl_df(data.frame(sampleData)))
#ggplot(sampleData_tbl, aes(x=value)) + geom_histogram()
#ggplot(sampleData_tbl, aes(x=value)) + geom_density()
```

### Sample Mean versus Theoretical Mean
```{r}
sampleMeans <- apply(sampleData, 1, mean)
sampleMean <- mean(sampleMeans)
sampleSD <- sd(sampleMeans)

inferredSD <- populationSD/sqrt(n)
```
```{r, echo=FALSE}
#hist(sampleMeans)
sampleMeans_tbl <- tbl_df(data.frame(sampleMeans))
ggplot(sampleMeans_tbl, aes(x=sampleMeans)) +
  geom_histogram(color="black", fill="white") +
  geom_vline(aes(xintercept=sampleMean), color="black", size=2) + 
  geom_vline(aes(xintercept=populationMean), color="red", linetype="dashed", size=1)
```

### Sample Variance versus Theoretical Variance
Compare the inferred variation of the means (inferredSD, via s/sqrt(n)) with the actual value
```{r, echo=FALSE}
#hist(sampleMeans)
ggplot(sampleMeans_tbl, aes(x=sampleMeans)) +
  geom_histogram(color="black", fill="white") +
  geom_vline(aes(xintercept=sampleMean), color="black", size=1, alpha=.2) + 
  geom_vline(aes(xintercept=populationMean), color="red", linetype="dashed", size=.5, alpha=.2) +
  geom_segment(aes(x=sampleMean-sampleSD, y=50, 
                   xend=sampleMean+sampleSD, yend=50), color="black", size=2) + 
  geom_segment(aes(x=populationMean-inferredSD, y=50, 
                   xend=populationMean+inferredSD, yend=50), color="red", linetype="dashed", size=1) +
  geom_segment(aes(x=sampleMean-2*sampleSD, y=25, 
                   xend=sampleMean+2*sampleSD, yend=25), color="black", size=2) + 
  geom_segment(aes(x=populationMean-2*inferredSD, y=25, 
                   xend=populationMean+2*inferredSD, yend=25), color="red", linetype="dashed", size=1)

```

### Distribution
tbd
```{r}
#x <- seq(0,50, length=100)
#hx <- dexp(x, rate=lambda)
```
```{r, echo=FALSE}
#plot(x, hx, type="l")
# http://stackoverflow.com/questions/6967664/ggplot2-histogram-with-normal-curve
ggplot(sampleMeans_tbl, aes(x=sampleMeans)) +
  geom_histogram(aes(y=..density..), color="black", fill="white") +
  #geom_density(alpha=.2, fill="#FF6666") +
  stat_function(fun=dnorm, args=list(mean=sampleMean, sd=sampleSD))
```

***





