---
title: 'Coursera Data Science Project: Statistical Inference (Part 2)'
author: "Rob Rolleston"
date: "August 23, 2015"
output: pdf_document
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=3, fig.path='figure/',
                      echo=TRUE, warning=FALSE, message=FALSE)
```
```{r echo=FALSE}
library(ggplot2, quietly = TRUE, verbose=FALSE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, verbose=FALSE, warn.conflicts = FALSE)
library(tidyr, quietly = TRUE, verbose=FALSE, warn.conflicts = FALSE)
```
###  Introduction
This is the project for the statistical inference class. In it, I will use simulation to explore inference and do some simple inferential data analysis. The project consists of two parts:

1. A simulation exercise.
2. Basic inferential data analysis (this report)


##  Basic Inferential Data Analysis

### Load and explore data
```{r}
library(datasets)
data("ToothGrowth")
ToothGrowth_tbl <- tbl_df(ToothGrowth)
glimpse(ToothGrowth_tbl)
```
Some inspection of the ToothGrowth data indicates it has `r nrow(ToothGrowth_tbl)` observations of `r ncol(ToothGrowth_tbl)` values: `r names(ToothGrowth_tbl)`.  The 'supp' value is a factor with only `r nlevels(as.factor(ToothGrowth_tbl$supp))` levels: `r levels(as.factor(ToothGrowth_tbl$supp))`. The 'dose' value, although a number, actually has only `r nlevels(as.factor(ToothGrowth_tbl$dose))` values: `r levels(as.factor(ToothGrowth_tbl$dose))`.  For processing, this value was converted to a factor.
```{r}
ToothGrowth_tbl$dose <- as.factor(ToothGrowth_tbl$dose)
```

### Data Summary
A faceted set of histograms
```{r}
ggplot(ToothGrowth_tbl, aes(x=len)) +
  geom_histogram(color="black", fill="white") +
  facet_grid(supp ~ dose) +
  ggtitle("Histograms of 'len' by 'supp' (OJ|VC) and 'dose' (0.5|1|2) ")
```

The basic question is: **"Is ToothGrowth 'len' affected by 'supp' or 'dose'?"**  
*Assumption: *  These observations are from a population of 60 individuals.  There are no paired samples.

The mean, sd, and count of values  of 'len' are:  
```{r}
summaryLen <- ToothGrowth_tbl %>% group_by(supp, dose) %>%
              summarize (mean = mean(len), sd=sd(len), n=n())
print(summaryLen)
```


### Compare tooth growth by supp and dose

Different doses of supp will be compared using a t-test.  To aid understanding about what is being tested, the figure below contains two sets of box plots; one for each of the the 'sup': 'OJ' and 'VC.  Each set contains 3 box plots, one for each of the doses: '0.5', '1', and '2'. The null hypothesis is: **There is no difference in the means of samples**
```{r}
supp_by_dose <- tbl_df(data.frame(matrix(ncol=4, nrow=6)))
colnames(supp_by_dose) <- c("supp", "dose1", "dose2", "pvalue")
supp_by_dose[1,] = c("OJ", ".5", "1", 
    t.test(len~dose, data=droplevels(filter(ToothGrowth_tbl, supp=="OJ" & dose!=2)))[3]$p.value)
supp_by_dose[2,] = c("OJ", ".5", "2", 
    t.test(len~dose, data=droplevels(filter(ToothGrowth_tbl, supp=="OJ" & dose!=1)))[3]$p.value)
supp_by_dose[3,] = c("OJ", "1", "2", 
    t.test(len~dose, data=droplevels(filter(ToothGrowth_tbl, supp=="OJ" & dose!=0.5)))[3]$p.value)

supp_by_dose[4,] = c("VC", ".5", "1", 
    t.test(len~dose, data=droplevels(filter(ToothGrowth_tbl, supp=="VC" & dose!=2)))[3]$p.value)
supp_by_dose[5,] = c("VC", ".5", "2", 
    t.test(len~dose, data=droplevels(filter(ToothGrowth_tbl, supp=="VC" & dose!=1)))[3]$p.value)
supp_by_dose[6,] = c("VC", "1", "2", 
    t.test(len~dose, data=droplevels(filter(ToothGrowth_tbl, supp=="VC" & dose!=0.5)))[3]$p.value)
supp_by_dose$pvalue <- round(as.numeric(supp_by_dose$pvalue), 4)
supp_by_dose <- mutate(supp_by_dose, H0 = ifelse(pvalue > 0.05, "Accept", "Reject"))
ggplot(ToothGrowth_tbl, aes(x=supp, y=len, fill=dose)) + geom_boxplot()
print(supp_by_dose)
```
As can be seen graphically, and confirmed by the p-values and rejection of the null hypothesis: *For each supplement, the amount of the 'dose' does indeed have a statistically different effect on the mean of the 'len'.*


Different 'supp' at each 'dose' will be compared using a t-test.  To aid understanding about what is being tested, the figure below contains three sets of box plots; one for each of the the 'dose': '0.5', '1', and '2'.  Each set contains 2 box plots, one for each of the 'sup' : 'OJ' and 'VC: . The null hypothesis is: **There is no difference in the means of samples**

```{r}
dose_by_supp <- tbl_df(data.frame(matrix(ncol=4, nrow=3)))
colnames(dose_by_supp) <- c("dose", "supp1", "supp2", "pvalue")
dose_by_supp[1,] <- c(".5", "OJ", "VC",
  t.test(len~supp, data=droplevels(filter(ToothGrowth_tbl,  dose==.5)))[3]$p.value)
dose_by_supp[2,] <- c("1", "OJ", "VC",
  t.test(len~supp, data=droplevels(filter(ToothGrowth_tbl,  dose==1)))[3]$p.value)
dose_by_supp[3,] <- c("2", "OJ", "VC",
  t.test(len~supp, data=droplevels(filter(ToothGrowth_tbl,  dose==2)))[3]$p.value)
dose_by_supp$pvalue <- round(as.numeric(dose_by_supp$pvalue), 4)
dose_by_supp <- mutate(dose_by_supp, H0 = ifelse(pvalue > 0.05, "Accept", "Reject"))
ggplot(ToothGrowth_tbl, aes(x=dose, y=len, fill=supp)) + geom_boxplot()
print(dose_by_supp)
```
As can be seen graphically, and confirmed by the p-values and rejection or acception of the null hypothesis: *For the 'doses' 0.5 and 1, the 'supp' does indeed have a statistically different effect on the mean of the 'len'.  for the dose of 2, there is no difference.*

### Conclusions
The only case where changing the supp or the dose does not affect the len is when the dose for each supp==2 (p-vale=0.96).  In all other cases the p-value is less than 0.05, and thus the null hypothesis is rejected.



