---
title: 'Coursera Data Science Project: Statistical Inference (Part 2)'
author: "Rob Rolleston"
date: "August 23, 2015"
output: pdf_document
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=3, fig.path='figure/',
                      echo=TRUE, warning=FALSE, message=FALSE)
```
```{r echo=FALSE}
library(ggplot2, quietly = TRUE, verbose=FALSE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, verbose=FALSE, warn.conflicts = FALSE)
library(tidyr, quietly = TRUE, verbose=FALSE, warn.conflicts = FALSE)
```
###  Introduction
This is the project for the statistical inference class. In it, I will use simulation to explore inference and do some simple inferential data analysis. The project consists of two parts:

1. A simulation exercise.
2. Basic inferential data analysis (this report)


##  Basic Inferential Data Analysis
tbd

### Load and explore data
```{r}
library(datasets)
data("ToothGrowth")
ToothGrowth_tbl <- tbl_df(ToothGrowth)
glimpse(ToothGrowth_tbl)
```
Some inspection of the ToothGrowth data indicates it has `r nrow(ToothGrowth_tbl)` observations of `r ncol(ToothGrowth_tbl)` values: `r names(ToothGrowth_tbl)`.  The 'supp' value is a factor with only `r nlevels(as.factor(ToothGrowth_tbl$supp))` levels: `r levels(as.factor(ToothGrowth_tbl$supp))`. The 'dose' value, although a number, actualy has only `r nlevels(as.factor(ToothGrowth_tbl$dose))` values: `r levels(as.factor(ToothGrowth_tbl$dose))`.  For processing, this value was converted to a factor.
```{r}
ToothGrowth_tbl$dose <- as.factor(ToothGrowth_tbl$dose)
```

### Data Summary
A faceted set of histograms
```{r}
ggplot(ToothGrowth_tbl, aes(x=len)) +
  geom_histogram(color="black", fill="white") +
  facet_grid(supp ~ dose) +
  ggtitle("Histograms of 'len' by 'supp' (OJ|VC) and 'dose' (0.5|1|2) ")
```

The basic question is: **"Is ToothGrowth 'len' afected by 'supp' or 'dose'?"**  
The mean, sd, and count of values  of 'len' are:  
```{r}
summaryLen <- ToothGrowth_tbl %>% group_by(supp, dose) %>%
              summarize (mean = mean(len), sd=sd(len), n=n())
print(summaryLen)
```


### Compare tooth growth by supp and dose
tbd

### Conclusions
tbd



